---
title: "Understanding the eradication of malaria in the United States 1920-1950"
subtitle: |
  Adam Howes (`ath19@ic.ac.uk`)
output:
  pdf_document:
    toc: false
    number_sections: true
    keep_tex: true
    includes:
      in_header: preamble.tex
bibliography: citations.bib
---

```{r echo = FALSE}
options(scipen = 100)
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  dpi = 320,
  cache = TRUE,
  out.width = "95%",
  fig.align = 'center'
)
comma <- function(x) formatC(x, format = "d", big.mark = ",")
```

# Background

* Malaria eradicated in the US during the years 1920-1950
* Interest in learning why this happened, or which factors were associated with faster eradication
* Have county level malaria data from various sources, with some limiations
* Have county level covariate data form various sources, with some limitations

# Data

* Malaria data as follows...
* Covariate data in the following categories
  * County
  * Drainage
  * Farmland
  * Mortality
  * People
  * Socioeconomic
  * Weather
  * Zooprophylaxis

# Model

* As implemented in `fit_sae-Y20-50`
* Uses `malariadata.csv` and `southern13_areas.geojson`
* Calculates mortality rate per 100,000
* For categorical data, takes some single fixed numeric value.
  * This could be improved using imputation via fitting a model, which I have started in `impute_numeric-Y33-37` but not finished
* Model defined for all years 1920 to 1950, and the counties from `southern13_areas.geojson`
* For covariates not available in a particular year and county, I imputed using `missForest()` and normalised the output (see `impute_covariates`)
  * I don't think the imputed covariates are good. Just looking at the time series output you get it's not what you'd want
* Model formula uses a fixed effect for each covariate, a Besag model on space at the country level, and an AR1 model on time at the year level
* Use a Poisson likelihood adapted to account for non-integer counts
* Fit the model with the `R-INLA` implementation of INLA

# Results

* Generate plausible seeming posterior means at a state level
* Don't particularly like the way that the data for 1933-1937 is handled
* Strange upswing at the end of the time period where there is no data
* Generate estimates of association with malaria rate of the covariates
  * Only association, no causation
  * Covariate inputation algorithm I don't think is good
  * Perhaps they're of interest anyway, I'm not sure

```{r fig.cap="Posterior mean of malaria rate at a state level as compared with the raw data."}
knitr::include_graphics("depends/time-series-state.pdf")
```

```{r fig.cap="Posterior mean and credible intervals for regression coefficient parameters for each covariate."}
knitr::include_graphics("depends/covariate-imporance.pdf")
```
